---
title: "Untitled"
author: "Saudat"
date: "25/05/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Install and load required packages
```{r, message=F, warning=F}
#BiocManager::install("clusterProfiler", version = "3.8")
#BiocManager::install("pathview")
#BiocManager::install("enrichplot")
#BiocManager::install("ggnewscale")
#BiocManager::install("europepmc")
BiocManager::install("pathview")

library(clusterProfiler)
library(enrichplot)
# we use ggplot2 to add x axis labels (ex: ridgeplot)
library(ggplot2)
library(pathview)

```

# Annotations
I'm using *D melanogaster* data, so I install and load the annotation "org.Dm.eg.db" below. See all annotations available here: http://bioconductor.org/packages/release/BiocViews.html#___OrgDb (there are 19 presently available). 
```{r, message=F, warning=F}
# SET THE DESIRED ORGANISM HERE
organism = "org.Dm.eg.db"
#BiocManager::install(organism, character.only = TRUE)
library(organism, character.only = TRUE)
organism = org.Dm.eg.db

```
#Prepare Input
```{r}
# reading in data from deseq2
df = read.csv("drosphila_example_de.csv", header=TRUE)
# we want the log2 fold change 
original_gene_list <- df$log2FoldChange
# name the vector
names(original_gene_list) <- df$X
# omit any NA values 
gene_list<-na.omit(original_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
```
## Gene Set Enrichment function
```{r}
gse <- gseGO(geneList=gene_list, 
             ont ="ALL", 
             keyType = "ENSEMBL", 
             nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none")
```
# Output
##Table of results
```{r}
head(gse)
```
##Dotplot
```{r echo=TRUE, fig.width=15, fig.height=8}
require(DOSE)
dotplot(gse, showCategory=10, split=".sign") + facet_grid(.~.sign)
```
##Encrichment plot map:
```{r echo=TRUE}
#emapplot(gse, showCategory = 10)
```
##Category Netplot
```{r fig.width=18}
# categorySize can be either 'pvalue' or 'geneNum'
cnetplot(gse, categorySize="pvalue", foldChange=gene_list, showCategory = 3)
```
## Ridgeplot
Helpful to interpret up/down-regulated pathways.
```{r fig.width=18, fig.height=12}
ridgeplot(gse) + labs(x = "enrichment distribution")
```
## GSEA Plot  
Traditional method for visualizing GSEA result.  
```{r fig.height=6}
# Use the `Gene Set` param for the index in the title, and as the value for geneSetId
gseaplot(gse, by = "all", title = gse$Description[1], geneSetID = 1)
```
## PubMed trend of enriched terms
Plots the number/proportion of publications trend based on the query result from PubMed Central.
```{r fig.width=10}
terms <- gse$Description[1:3]
pmcplot(terms, 2010:2018, proportion=FALSE)
```
# KEGG Gene Set Enrichment Analysis
## Prepare Input
```{r}
# Convert gene IDs for gseKEGG function
# We will lose some genes here because not all IDs will be converted
ids<-bitr(names(original_gene_list), fromType = "ENSEMBL", toType = "ENTREZID", OrgDb=organism)
# remove duplicate IDS (here I use "ENSEMBL", but it should be whatever was selected as keyType)
dedup_ids = ids[!duplicated(ids[c("ENSEMBL")]),]
# Create a new dataframe df2 which has only the genes which were successfully mapped using the bitr function above
df2 = df[df$X %in% dedup_ids$ENSEMBL,]
# Create a new column in df2 with the corresponding ENTREZ IDs
df2$Y = dedup_ids$ENTREZID
# Create a vector of the gene unuiverse
kegg_gene_list <- df2$log2FoldChange
# Name vector with ENTREZ ids
names(kegg_gene_list) <- df2$Y
# omit any NA values 
kegg_gene_list<-na.omit(kegg_gene_list)
# sort the list in decreasing order (required for clusterProfiler)
kegg_gene_list = sort(kegg_gene_list, decreasing = TRUE)
```

```{r}
kegg_organism = "dme"
kk2 <- gseKEGG(geneList     = kegg_gene_list,
               organism     = kegg_organism,
               nPerm        = 10000,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "ncbi-geneid")
```

```{r}
head(kk2, 10)
```
## Dotplot
```{r echo=TRUE}
dotplot(kk2, showCategory = 10, title = "Enriched Pathways" , split=".sign") + facet_grid(.~.sign)
```
## Encrichment plot map:
```{r echo=TRUE}
#emapplot(kk2)
```
## Category Netplot:
```{r fig.width=12}
# categorySize can be either 'pvalue' or 'geneNum'
cnetplot(kk2, categorySize="pvalue", foldChange=gene_list)
```
## Ridgeplot
Helpful to interpret up/down-regulated pathways.
```{r fig.width=18, fig.height=12}
ridgeplot(kk2) + labs(x = "enrichment distribution")
```
# GSEA Plot  
Traditional method for visualizing GSEA result.  
```{r fig.height=6}
# Use the `Gene Set` param for the index in the title, and as the value for geneSetId
gseaplot(kk2, by = "all", title = kk2$Description[1], geneSetID = 1)
```
#Pathview
```{r, message=F, warning=F, fig.height=6}
# Produce the native KEGG plot (PNG)
dme <- pathview(gene.data=kegg_gene_list, pathway.id="dme04130", species = kegg_organism)
# Produce a different plot (PDF) (not displayed here)
dme <- pathview(gene.data=kegg_gene_list, pathway.id="dme04130", species = kegg_organism, kegg.native = F)
```






